Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-07-06 PAPER.tex
-Re-organized layout of paper so that themes are discussed first, with PAPER-64 application at the end -Fleshed out themes intro -Fleshed out signal loss toy model section

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of the omnipresent neutral hydrogen. This important epoch of our cosmic dawn harbors a wealth of information regarding the environment during this transformative time, including insight into the nature of the first luminous sources and implications about cosmological parameters. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. As instruments become more sensitive and a statistical power spectrum detection is in our foreseeable future, it has become increasingly important to develop techniques that help ensure that a potential detection isn't washed out or mistaken for something else. In this paper, we focus on two crucial ingredients for a power spectrum detection - a method for the accurate injection and recovery of an EoR signal, and jack-knife tests that can tease out foregrounds and systematics that contaminate our signal. We describe and apply these methods using data taken by the 64-element configuration of the Donald C. Baker Precision Array for Probing the Epoch of Reionization (PAPER).
\end{abstract}

@@

The major challenge that faces all 21 cm experiments is in isolating a small signal that is buried underneath foregrounds and instrumental systematics that are 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a very careful, thorough understanding of data analysis choices. With HERA on the horizon and continual progress being made in the field, it is becoming increasingly more important to develop methods that can both reinforce a potential real detection and explain any false ones. In this paper, we present approaches to both of these issues using data from the 64-element configuration of PAPER. We place special emphasis on the accurate injection and recovery of a fake EoR signal in order to assess signal loss associated with different weighting techniques, as well as the importance of using jack-knife tests to investigate sources of bias.

This paper is organized as follows. In section 2 we give a brief overview of the PAPER-64 array and observations. In section 3 we describe our power spectrum estimation pipeline. Section 4 presents two key components of any power spectrum analysis pipeline - accurately quantifying signal loss that may result from data analysis choices, and choosing a weighting scheme to maximum foreground down-weighting without amassing large amounts of signal loss. In Section 5 we highlight useful jack-knife tests that can help discriminate between excess detections from foregrounds. We conclude in Section 6.

\section{Observations}
\label{sec:Obs}

@@

\begin{figure*}
	\centering
	\includegraphics[height=0.35\textwidth]{paper_dipole.png}
	\includegraphics[height=0.35\textwidth]{antenna_layout.png}
	\caption{PAPER dipole in South Africa (left) and PAPER-64 antenna layout (right).}
	\label{fig:paper}
\end{figure*}

@@

Finally, after removing an additional layer of RFI (values $3\sigma$ above the median), we average together all our data into two datasets: even julian dates and odd julian dates. A total of $124$ nights of data comprises this average, and we use these two datasets as a starting point for constructing power spectrum estimates.

\section{Power Spectrum Analysis}
\label{sec:PS}

\subsection{Fringe-Rate-Filtering}
\label{sec:FRF}

As shown in \citet{parsons_et_al2016}, a fringe-rate filter allows us to maximize our sensitivity even further by averaging visibilities in time. We apply this filter in the fringe-rate domain, the Fourier-dual to time, and it effectively amounts to increasing our data's integration time. Another way to describe the filter is that it weights different parts of the sky (which rotate at different ``fringe-rates") by the antenna beam. This allows us to up-weight parts of the sky high in the primary beam, and down-weight those that are less sensitive.

In \citet{ali_et_al2015}, the filter applied was degraded (widened in fringe-rate space) from an optimal one that can be constructed based on the integration time of the data. This was chosen to decrease the resulting integration time, therefore increasing the number of independent modes and reducing signal loss associated with applying optimal quadratic estimators (OQE) on data with too few modes.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{frp.png}
	\caption{The optimal fringe-rate filter used in the analysis, normalized to $1$.}
	\label{fig:frp}
\end{figure}

In practice, the optimal fringe-rate profile is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter is implemented on a per baseline basis by convolving the time-domain visibilities with the Fourier transform of the fringe-rate filter, resulting in averaged visibilities yielding $\sim 2$ orders of magnitude in sensitivity.

\subsection{Power Spectrum Estimation}
\label{sec:PSEst}

@@

We begin with our data vector $\textbf{x}$, which contains our visibilities that have the shape ($times$, $frequencies$) for each baseline and data group (`even' or `odd'). As part of our bootstrapping routine (and to speed things up), we divide up our total number of baselines into $5$ random, equal-sized groups, where no baselines are repeated within or between groups. For the \citet{ali_et_al2015} method, each group is then sampled with replacement to create a new group, of the same size, that can have repeated baselines inside it. We discover that in doing so, we are sacrificing some of our sensitivity since this results in there being 3-4 repeated baselines per group. In order to maximize our sensitivity but still apply random sampling for use in error estimation, we instead form new groups using all independent baselines except the very last one. For example, if we have $10$ baselines in a group, we use the first $9$ to guarantee at least $9$ independent measurements, and then fill the last slot randomly out of the $10$. We do this for all $5$ groups. This is still a valid means of bootstrapping because there are many more possibilities of baseline groupings than the number of bootstraps we run for this analysis ($nboots = 20$).

Taking $\textbf{x}$ to be averaged data from all baselines within a group, we now have $10$ possible data vectors (for $5$ groups and $2$ datasets). We next form the quantity $\hat{q}_{\alpha}$:

@@

The rest of the OQE formalism is identical to \citet{ali_et_al2015}. We normalize our power spectrum estimates using the matrix $\textbf{M}$:

@@

In practice, we end up with a 2-dimensional power spectrum quantity, which is a function of both time and $k$. We compute $20$ bootstraps of these quantities, where each bootstrap creates baseline groups randomly.

In \citet{ali_et_al2015}, a second round of bootstrapping occurs over the bootstrap and time axes simultaneously. Random values are sampled with replacement along both axes, drawing as many values as there are number of bootstraps and times. Final power spectrum limits are then computed by taking the mean and standard deviation over this second bootstrap axis.

In summary, our power spectrum pipeline begins with data with four axes (datasets, baselines, times, frequencies), bootstraps over baselines, uses OQE formalism, and produces $1$-dimensional power spectra for both $\textbf{P(k)}$ and the folded version $\Delta^{2}\textbf{(k)}$, defined as:

@@

@@

\subsection{Power Spectrum Sensitivity}
\label{sec:PSSense}

\section{Signal Loss}
\label{sec:Sigloss}

\subsection{Formalism}
\label{sec:Formalism}

@@

\subsection{Application}
\label{sec:siglossprac}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{sigloss1_noise.png}
	\caption{$P_{in}$ vs. $P_{out}$ (blue points) for 15 injection levels and 21 $k$'s. The solid cyan line is the $2\sigma$ upper limit for the weighted power spectrum of noise alone, and it is at this level where the signal loss factor $P_{in}/P_{out}$ is computed by interpolation. \cc{Maybe re-do this plot with closer-together points}}
	\label{fig:sigloss1_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{ps1_noise_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{ps1_noise.png}
	\caption{Full inverse covariance weighted power spectrum of pure noise (black points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters. \cc{Color negative points grey}}
	\label{fig:ps1_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{sigloss2_noise.png}
	\caption{$P_{in}$ vs. $P_{out}$ (blue) for 15 injection levels, 20 bootstraps, and 21 $k$'s. \cc{Plot a semi-transparent cyan range of pCn values instead of just the max}}
	\label{fig:sigloss2_noise}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{ps2_noise_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{ps2_noise.png}
	\caption{Full inverse covariance weighted power spectrum of pure noise (black and grey points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). Black points correspond to positive values, while grey points correspond to originally negative values that have been made positive for plotting. The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters.}
	\label{fig:ps2_noise}
\end{figure*}

@@

%\begin{figure*}
%	\centering
%	\includegraphics[width=1.0\textwidth]{sigloss1_data.png}
%	\caption{$P_{in}$ vs. $P_{out}$ (black) for 15 injection levels and 21 $k$'s. The solid grey line is the $2\sigma$ upper limit for the weighted power spectrum of data alone, and it is at this level where the signal loss factor $P_{in}/P_{out}$ is computed by interpolation. \cc{Probably don't need to show this plot at all}}
%	\label{fig:sigloss1_data}
%\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{sigloss2_data.png}
	\caption{$P_{in}$ vs. $P_{out}$ (black) for 15 injection levels, 20 bootstraps, and 21 $k$'s. \cc{Plot a semi-transparent grey range of pCv values instead of just the max}}
	\label{fig:sigloss2_data}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{ps2_data_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{ps2_data.png}
	\caption{Full inverse covariance weighted power spectrum of PAPER-64 data (black and grey points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). Black points correspond to positive values, while grey points correspond to originally negative values that have been made positive for plotting. The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{eigenspectrum.png}
	\includegraphics[width=0.4\textwidth]{covariance.png}
	\caption{Eigenspectrum for $\textbf{C}$ for baseline (1,4) for the 21 channels ofs interest (left) and covariance matrix $\textbf{C}$ for the same baseline (right).}
	\label{fig:eigenspectrum}
\end{figure*}

@@

\section{Jack-Knives}
\label{sec:Jack}

>>>>

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen. This important epoch of our cosmic dawn harbors a wealth of information regarding the environment during this transformative time, including insight into the nature of the first luminous sources and implications about cosmological parameters. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes - signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Baker Precision Array for Probing the Epoch of Reionization (PAPER).
\end{abstract}

@@

The major challenge that faces all 21 cm experiments is in isolating a small signal that is buried underneath foregrounds and instrumental systematics that are 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a very careful, thorough understanding of data analysis choices. With HERA on the horizon and continual progress being made in the field, it is becoming increasingly more important to understand how the methods we choose affect each other and any power spectrum results. In this paper, we discuss three themes essential to a $21$ cm power spectrum study and how data analysis choices affect each. We approach these themes from both a broad perspective, and through a detailed case study using data from the 64-element configuration of PAPER.

This paper is organized as follows. In section 2 we introduce the three themes of our focus, using a toy model to develop intuition into each one. Section 3 presents an overview of the PAPER-64 array and observations, highlighting key changes from \citet{ali_et_al2015}. Sections 4, 5, and 6 detail how the new PAPER-64 analysis quantifies signal loss, estimates error bars, and eliminates bias, respectively. We conclude in Section 7.

\section{Power Spectrum Themes and Techniques}
\label{sec:Themes}

A typical dataset for a $21$ cm experiment involves many axes. Visibility measurements, the standard data product for radio telescopes, are saved for many baselines and baseline-types. These visibilities are comprised of many integrations and a spectrum of frequency channels. Data collection can span months and years and can involve multiple polarizations. The data itself contains not only an EoR signal, but layers of bright foregrounds, radio frequency interference, instrumental noise, and other systematics.

The job for $21$ cm data analysts is to sift through terabytes of data to ultimately make a power spectrum, a statistical result that encodes the variance of a dataset. In other words, a power spectrum measures the strength of the $21$ cm signal on various spatial scales of the sky, a powerful measurement that can be used to trace the evolution of neutral hydrogen during the EoR. There are, however, many data analysis choices that affect both the variance of the dataset itself and how it is measured.

Simply stated, we desire to maximize the sensitivity of a dataset as much as possible, while being confident about the result and able to account for irregularities. There are many choices as a data analyst, such as how to optimally combine time-ordered measurements, how to best and most accurately estimate its variance, and how to weight data in a way that suppresses uninteresting modes while not destroying an EoR signal. Common techniques such as averaging data, inverse covariance weighting, bootstrapping, and jack-knife testing each affect data in different ways and reveal different lessons.

There are often tradeoffs between techniques. For example, \cc{cite papers such as Josh's}.

In this paper, we focus on four power spectrum techniques and their affect on three umbrella $21$ cm power spectrum themes. We will give brief definitions now, and build intuition for each theme in the sections to follow.

\begin{center}
Power Spectrum Techniques
\end{center}
\begin{itemize}
\item \textbf{Fringe-rate filtering:} The concept of a fringe-rate filter is similar to averaging data in time. We explain our filter in more detail in Section \ref{sec:CaseStudy}, but a broad description is that a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window.
\item \textbf{Inverse covariance weighting:} Of the same flavor as inverse variance weighting, inverse covariance weighting weights a dataset by minimizing the covariance between frequency channels. This weighting has the effect of down-weighting correlated information and up-weighting noise-like information.
\item \textbf{Bootstrapping:} This is a useful method for estimating errors of a dataset from itself. By randomly drawing samples of the data, the spread in those numbers gives us our errors.
\item \textbf{Jack-Knife testing:} A resampling technique useful for estimating bias, jack-knives can be taken along different dimensions of a dataset to cross-validate results.
\end{itemize}

\begin{center}
Power Spectrum Themes
\end{center}
\begin{itemize}
\item \textbf{Signal Loss:} As explained in the next session, there are analysis techniques that can lead to the loss of an EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss has subtle challenges but is a crucial component for confidence in any result.
\item \textbf{Error Bar Estimation:} Errors on a $21$ cm power spectrum result can make the difference between a detection and a noise-like measurement, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias:} There are several possible sources of bias in a visibility measurement that can leak its way into a power spectrum. It is important to identity them in order to interpret results and develop techniques to suppress contamination.
\end{itemize}

We will now discuss the three themes more in depth and discuss how certain power spectrum techniques affect each.

\subsection{Signal Loss}
\label{sec:SiglossOverview}

Signal loss refers to the loss of a cosmological signal in a dataset. It can arise in a variety of ways at multiple stages in an analysis pipeline. Here we focus on signal loss originating from empirically estimating covariance matrices, a method that succeeds in down-weighting foregrounds but can carry the risk of large amounts of signal loss. We will develop our intuition and motivate the effect of signal loss using a toy model simulation.

Suppose we have a very simple dataset which contains 2-dimensional data (representing $100$ time integrations and $20$ frequency channels). The data is comprised of a sine-wave foreground that varies in time, and a complex, random-noise EoR signal (Figure \ref{fig:toy_sigloss1}, left).

\begin{figure*}
	\centering
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss1.png}
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss2.png}
	\caption{Left: Our toy model dataset contains a sine-wave foreground that varies in time and random noise as an EoR signal. Real parts are shown here. Right: Eigenspectrum of foregrounds only (blue), EoR only (red), and both (green)}
	\label{fig:toy_sigloss1}
\end{figure*}

Before we demonstrate how signal loss can arise, we first look at the eigenspectrum of the covariance matrix that describes our data. That is, if $\textbf{x}$ is our dataset which contains both the foreground and EoR, its covariance matrix is empirically defined as $\textbf{C} \equiv \textbf{xx}^{\dagger}$. The eigenspectrum, which encodes useful information about the strengths of different frequency modes, is shown in the right plot of Figure \ref{fig:toy_sigloss1}. It is obvious that there is only 1 eigenmode in the foreground-only dataset (since we used one sine-wave), and therefore most of the information in our combined dataset is contained in the first mode.

We would like to use inverse covariance weighting to down-weight the foregrounds in our dataset. To do so, we compute $\textbf{C}^{-1}$ and use optimal quadratic estimators (see Section \ref{sec:PS} for more details) to Fourier-transform our data along the `frequency' axis and form a power spectrum. As an explicit check, we use information from only the first eigenmode, and zero out the others. We do this because we know that the foregrounds can be perfectly described by only the first mode, so it should be all that we need to down-weight them.

Figure \ref{fig:toy_sigloss3} shows the unweighted (using the identity matrix $\textbf{I}$ instead of $\textbf{C}$) power spectrum for the foreground-only data in blue, the unweighted power spectrum for the EoR-only data in red, and the weighted power spectrum for the combined dataset in green. There is one peak in blue from our one foreground mode.

\begin{figure*}
	\centering
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss4.png}
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss3.png}
	\caption{Power spectrum of foregrounds only (blue), EoR only (red), and the combined dataset (green). Contrasted are the effect of using information from the first eigenmode only (left), and full inverse covariance weighting (right).}
	\label{fig:toy_sigloss3}
\end{figure*}

From the left plot, we see that by down-weighting the first mode, we can perfectly recover EoR. This is in principle the reason why inverse covariance weighting is so attractive for $21$ cm data. If we could perfectly model our foregrounds, we could down-weight them without any signal loss. However, this is not the case in reality, since we can't distinguish between foreground and EoR modes.

If we did not know that our foreground information is contained in only one mode, a naive approach would be to use full inverse covariance weighting (using information from all eigenmodes). Examining this case, our full inverse covariance weighted result (Figure \ref{fig:toy_sigloss3}, right) almost recovers the EoR signal. It exhibits the correct shape, but the amplitude level is slightly low. This is actually evidence of a small amount of signal loss. In this toy model, the amount of signal loss is small, but we can exaggerate the effect and investigate its cause by `fringe-rate filtering' our toy model dataset.

Since fringe-rate filtering is similar in practice to averaging in time, we average every $4$ time integrations together, yielding $25$ independent samples in time (Figure \ref{fig:toy_sigloss5}, left). We choose these numbers so that the total number of independent samples is similar to the number of frequency channels (our matrices will be full rank). The resulting eigenspectra (Figure \ref{fig:toy_sigloss5}, right), as compared to those in Figure \ref{fig:toy_sigloss1}, begin to fall for the last few eigenmodes. Although a subtle feature, it has large consequences on the resulting power spectra.

\begin{figure*}
	\centering
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[height=0.3\textwidth]{plots/toy_sigloss6.png}
	\caption{Left: Our `fringe-rate filtered' toy model dataset. It contains $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of foregrounds only (blue), EoR only (red), and both (green).}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Power spectrum of foregrounds only (blue), EoR only (red), and the full inverse covariance weighted combined dataset (green) for a time-averaged dataset (fewer independent modes).}
	\label{fig:toy_sigloss7}
\end{figure}

Applying full inverse covariance weighting to this dataset results in the power spectra shown in Figure \ref{fig:toy_sigloss7}. It is clear that there is a much larger amount of signal loss for this time-averaged dataset. The subtle drop-off in the eigenspectrum shape suggests that when the spectrum is inverted, a few modes are up-weighted more strongly than the rest. This means that most of our power spectrum information is coming from only the last few modes, and as a result, is a noisier estimate. This is evident by noticing that the green curve in Figure \ref{fig:toy_sigloss7} not only exhibits signal loss, but also fails to trace the shape of the unweighted EoR power spectrum.

Using our toy model, we have shown that signal loss can arise from inverse covariance weighting due to unintentionally up-weighting eigenmodes with very small eigenvalues (overfitting the noise). We have also shown that this effect is made larger by reducing the total number of independent samples in a dataset. In reality, $21$ cm data from PAPER-64 exhibits $\sim3$ orders of magnitude of signal loss when combining the use of an optimal fringe-rate filter and full inverse covariance weighting.

Going back to our time-averaged simulation data, we will now describe several ways to alter the covariance matrix $\textbf{C}$ in order to avoid signal loss. Recall that $\textbf{C}$ is computed from the dataset itself, leading to the up-weighting of very noisy, non-existent modes. In practice we compute $\textbf{C}$ this way because we do not know how to accurately model the matrix. But for our toy model, we know that our simulated, perfectly-noisy EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such and add it to $\textbf{C}_{FG} = \textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}$ to obtain a final $\textbf{C}$ to use in weighting (instead of computing it based on $(\textbf{x}_{FG} + \textbf{x}_{EoR})$), we see that there is no signal loss (Figure \ref{fig:toy_sigloss8}, left).

\begin{figure*}
	\centering
	\includegraphics[height=0.21\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[height=0.21\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[height=0.21\textwidth]{plots/toy_sigloss9.png}
	\caption{Power spectra of foregrounds only (blue), EoR only (red), and the weighted combined dataset (green). We show three alternate weighting options that each avoid signal loss, including modeling the covariance matrix of EoR (left), regularizing $\textbf{C}$ by adding an identity matrix to it (middle), and using only the first few eigenmodes of $\textbf{C}$ (right).}
	\label{fig:toy_sigloss8}
\end{figure*}

\cc{add paragraph about regularization options}

\subsection{Error Estimation}
\label{sec:ErrorOverview}

\subsection{Bias}
\label{sec:BiasOverview}

\section{Case Study: PAPER-64}
\label{sec:CaseStudy}

Now that we have developed our intuition regarding signal loss, error estimation, and bias, we will look at each in more detail as applied to data taken by PAPER. We begin with an overview of the telescope array, its observations, and analysis pipeline.

\subsection{Observations}
\label{sec:Obs}

@@

\begin{figure*}
	\centering
	\includegraphics[height=0.35\textwidth]{plots/paper_dipole.png}
	\includegraphics[height=0.35\textwidth]{plots/antenna_layout.png}
	\caption{PAPER dipole in South Africa (left) and PAPER-64 antenna layout (right).}
	\label{fig:paper}
\end{figure*}

@@

After removing an additional layer of RFI (values $3\sigma$ above the median), we average together all our data into two datasets: even julian dates and odd julian dates. A total of $124$ nights of data comprises this average.

Finally, our last analysis step before power spectrum estimation is the use of a fringe-rate filter. As shown in \citet{parsons_et_al2016}, a fringe-rate filter allows us to maximize our sensitivity even further by averaging visibilities in time. We apply this filter in the fringe-rate domain, the Fourier-dual to time, and it effectively amounts to increasing our data's integration time. Another way to describe the filter is that it weights different parts of the sky (which rotate at different ``fringe-rates") by the antenna beam. This allows us to up-weight parts of the sky high in the primary beam, and down-weight those that are less sensitive. In practice, the optimal fringe-rate profile is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter is implemented on a per baseline basis by convolving the time-domain visibilities with the Fourier transform of the fringe-rate filter, resulting in averaged visibilities yielding $\sim 2$ orders of magnitude in sensitivity.

\subsection{Power Spectrum Estimation}
\label{sec:PS}

@@

We begin with our data vector $\textbf{x}$, which contains our visibilities that have the shape ($times$, $frequencies$) for each baseline and data group (`even' or `odd'). As part of our bootstrapping routine (and to speed things up), we divide up our total number of baselines into $5$ random, equal-sized groups. Taking $\textbf{x}$ to be averaged data from all baselines within a group, we now have $10$ possible data vectors (for $5$ groups and $2$ datasets). We next form the quantity $\hat{q}_{\alpha}$:

@@

Finally, we normalize our power spectrum estimates using the matrix $\textbf{M}$:

@@

In practice, we end up with a 2-dimensional power spectrum quantity, which is a function of both time and $k$. We compute $20$ bootstraps of these quantities, where each bootstrap creates baseline groups randomly. We form both the quantity $\textbf{P(k)}$ and the folded version $\Delta^{2}\textbf{(k)}$, defined as:

@@

\section{Application: Signal Loss}
\label{sec:Sigloss}

\subsection{Noise and EoR}

@@

\subsection{Signal Loss Methodology}

@@

\subsection{Signal Loss in Practice}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{plots/sigloss1_noise.png}
	\caption{$P_{in}$ vs. $P_{out}$ (blue points) for 15 injection levels and 21 $k$'s. The solid cyan line is the $2\sigma$ upper limit for the weighted power spectrum of noise alone, and it is at this level where the signal loss factor $P_{in}/P_{out}$ is computed by interpolation. \cc{Maybe re-do this plot with closer-together points}}
	\label{fig:sigloss1_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise.png}
	\caption{Full inverse covariance weighted power spectrum of pure noise (black points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters. \cc{Color negative points grey}}
	\label{fig:ps1_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{plots/sigloss2_noise.png}
	\caption{$P_{in}$ vs. $P_{out}$ (blue) for 15 injection levels, 20 bootstraps, and 21 $k$'s. \cc{Plot a semi-transparent cyan range of pCn values instead of just the max}}
	\label{fig:sigloss2_noise}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise.png}
	\caption{Full inverse covariance weighted power spectrum of pure noise (black and grey points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). Black points correspond to positive values, while grey points correspond to originally negative values that have been made positive for plotting. The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters.}
	\label{fig:ps2_noise}
\end{figure*}

@@

%\begin{figure*}
%	\centering
%	\includegraphics[width=1.0\textwidth]{plots/sigloss1_data.png}
%	\caption{$P_{in}$ vs. $P_{out}$ (black) for 15 injection levels and 21 $k$'s. The solid grey line is the $2\sigma$ upper limit for the weighted power spectrum of data alone, and it is at this level where the signal loss factor $P_{in}/P_{out}$ is computed by interpolation. \cc{Probably don't need to show this plot at all}}
%	\label{fig:sigloss1_data}
%\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{plots/sigloss2_data.png}
	\caption{$P_{in}$ vs. $P_{out}$ (black) for 15 injection levels, 20 bootstraps, and 21 $k$'s. \cc{Plot a semi-transparent grey range of pCv values instead of just the max}}
	\label{fig:sigloss2_data}
\end{figure*}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_nosigloss.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.png}
	\caption{Full inverse covariance weighted power spectrum of PAPER-64 data (black and grey points, with $2\sigma$ error bars) before signal loss correction (left) and after (right). Black points correspond to positive values, while grey points correspond to originally negative values that have been made positive for plotting. The dashed blue line is the unweighted power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical noise level prediction based on observational parameters.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/eigenspectrum.png}
	\includegraphics[width=0.4\textwidth]{plots/covariance.png}
	\caption{Eigenspectrum for $\textbf{C}$ for baseline (1,4) for the 21 channels ofs interest (left) and covariance matrix $\textbf{C}$ for the same baseline (right).}
	\label{fig:eigenspectrum}
\end{figure*}

@@

\section{Application: Error Estimation}
\label{sec:Error}

\subsection{Thermal Noise}

\subsection{Bootstrapping}

\cc{Include info about baseline bootstrapping:}
For the \citet{ali_et_al2015} method, each group is then sampled with replacement to create a new group, of the same size, that can have repeated baselines inside it. We discover that in doing so, we are sacrificing some of our sensitivity since this results in there being 3-4 repeated baselines per group. In order to maximize our sensitivity but still apply random sampling for use in error estimation, we instead form new groups using all independent baselines except the very last one. For example, if we have $10$ baselines in a group, we use the first $9$ to guarantee at least $9$ independent measurements, and then fill the last slot randomly out of the $10$. We do this for all $5$ groups. This is still a valid means of bootstrapping because there are many more possibilities of baseline groupings than the number of bootstraps we run for this analysis ($nboots = 20$).

\cc{Include info about second round of bootstrapping:}
In \citet{ali_et_al2015}, a second round of bootstrapping occurs over the bootstrap and time axes simultaneously. Random values are sampled with replacement along both axes, drawing as many values as there are number of bootstraps and times. Final power spectrum limits are then computed by taking the mean and standard deviation over this second bootstrap axis.

\section{Application: Bias}
\label{sec:Bias}

\cc{Include info about optimal FRF:}
In \citet{ali_et_al2015}, the filter applied was degraded (widened in fringe-rate space) from an optimal one that can be constructed based on the integration time of the data. This was chosen to decrease the resulting integration time, therefore increasing the number of independent modes and reducing signal loss associated with applying optimal quadratic estimators (OQE) on data with too few modes.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/frp.png}
	\caption{The optimal fringe-rate filter used in the analysis, normalized to $1$.}
	\label{fig:frp}
\end{figure}
